Comparative Evaluation of Spark and Flink Stream Processing

### 2017-03-07 paper/paper.tex
[paper] updates

  \centering
    \includegraphics[width=\textwidth, height=.3\textheight]{flink_engine.png}
     \caption{ An example of data flow graph in Flink \cite{flink}.}
\end{figure} 
\subsection{Apache Kafka}
\par Apache Kafka is scalable, fault-tolerant distributed publish/subscribe streaming framework \cite{kafka}.
It manges the stream records in different categorizes (i.e., topics) that portioned and distributed over the servers of the Kafka cluster. It allows the data producers to publish stream of records to certain Kafka topic. While the consumer applications can subscribe to one or more topic to read the data stream. The consumers are grouped into groups based on the group name that allows Kafka to distribute and balance the stream partitions among the  members of a certain group for the sake of scalability. Figure 3 shows how 4 partitions of stream is distributed differentially between two groups of consumers based on the number of members. 
Apache Kafka has been widely  adopted, for example, 
Spark and Flink can receive data stream from Kafka.

@@

  \centering
    \includegraphics[width=\textwidth, height=.3\textheight]{kafka_groups.png}
     \caption{ Distribution of stream partitions for consumer groups in Kafka \cite{kafka}.}
\end{figure}

\section{Experiment Setup and Implementation}
This section describes the data stream setup, the design of evaluation streaming workloads, and the implementation details in Spark and Flink.

\subsection{Data Stream Setup}
The evaluation streaming workloads read an input data stream from Kafka, in order to simulate real-world use case we use datasets of Automatic Dependent Surveillance – Broadcast (ADS-B) messages that present the aircraft's position over time, comprise 22 fields of data such as aircraft ID, date message generated, longitude, latitude, and altitude etc.
\par In our experiment a data stream producer component reads the ADS-B message datasets, then publish them to Kafka cluster to be consumed by the workloads in Apache Spark and Apache Flink. Figure 4 illustrates the data producer and the Kafka cluster setup in all experiments.
The data stream is portioned into four portions over two server, while the data producer publish the stream records randomly to Kafka partitions, and all ADS-B messages are published to same topic, which is \texttt{datacorn}. Moreover, the data producer attach the streaming time at the ADS-B message line before publishing it to the Kafka topic, to be used in the latency calculation.

@@

  \centering
    \includegraphics[width=.9\textwidth, height=.2\textheight]{kafka.png}
     \caption{Data Stream Producer \& Kafka Cluster Setup.}
\end{figure}

\subsection{Evaluation Streaming Workloads}
In our experimental evaluation, we developed two real-time stream processing workloads that read the ADS-B messages stream from Kafka in order to perform basic trajectories analysis methods. The workloads are designed in such way that cover general key aspects of real-time streaming processing tasks, and evaluate the corresponding solutions in Spark and Flink. The following are the aspects of streaming data processing were covered by our workloads's design:

\begin{itemize}
\item Handling parallel input streams (e.g., Kafka Stream).
\item How to aggregate the state of input stream.
\item Manage the order of stream records.
\item How to provide and update global data model in stream processing task.
\item Evaluate the performance by measuring  the latency and throughput.
\end{itemize}
We present the description of workloads and the relation with these aspects, also the implementation details of Spark and Flink solutions in Section 3.2.1 \& 3.2.1. Afterward, In Section 4 we discuss the performance evaluation and analysis of these workloads in both frameworks.

>>>>

  \centering
    \includegraphics[width=\textwidth, height=.3\textheight]{flink_engine.png}
     \caption{ An example of data flow graph in Flink \cite{flink}.}
\end{figure} 
\subsection{Apache Kafka}
\par Apache Kafka is scalable, fault-tolerant and distributed streaming framework \cite{kafka}. It allows to publish and subscribe to data streams.
Kafka manages the stream records in different categories (i.e., topics) that partitioned and distributed over the servers of the Kafka cluster. It allows the data producers to publish a stream of records to one or more  Kafka topic. Additionally, consumer applications can subscribe to one or more topic to read data streams. The consumers are grouped into groups based on the group name that allows Kafka to distribute and balance the stream partitions among the  members of a certain group for the sake of scalability. Figure 3 shows how 4 partitions of a stream are distributed between two groups of consumers based on the number of members. 
Apache Kafka has been widely  adopted, for example, 
Apache Spark and Apache Flink can ingest Kafka streams.

@@

  \centering
    \includegraphics[width=\textwidth, height=.3\textheight]{kafka_groups.png}
     \caption{ Distribution of a stream partitions for consumer groups in Kafka \cite{kafka}.}
\end{figure}

\section{Experiment Setup and Implementation}
This section describes the data stream setup, the design of evaluation streaming workloads and the solution's implementation details of  Apache Spark and Apache Flink.

\subsection{Data Stream Setup}
The evaluation streaming workloads read an input data stream from Kafka. In order to simulate real-world use case we use datasets of Automatic Dependent Surveillance – Broadcast (ADS-B) messages that present the aircraft's position over time. The message line comprises 22 fields of data such as aircraft ID, date message generated, longitude, latitude and altitude.
\par In our experiment a data stream producer component reads the ADS-B message datasets, then publishes them to the Kafka cluster to be consumed by the workloads in Apache Spark and Apache Flink. Figure 4 illustrates the data producer and the Kafka cluster setup in all experiments.
The data stream is divided into four portions over two server. The data producer publishes the stream records randomly to a Kafka partition and all ADS-B messages are published to the same topic (\texttt{datacorn}). Moreover, the data producer attaches the time of publishing to the ADS-B message line before publishing it to the Kafka topic.

@@

  \centering
    \includegraphics[width=.9\textwidth, height=.2\textheight]{kafka.png}
     \caption{The setuop of the Data Stream Producer \& Kafka Cluster.}
\end{figure}

\subsection{Evaluation Streaming Workloads}
In our experimental evaluation, we developed two real-time stream processing workloads that read a ADS-B messages stream from Kafka in order to perform basic trajectories analysis methods. The workloads are designed to cover general key aspects of real-time stream processing tasks and evaluate the corresponding solutions in Spark and Flink. The following are the aspects of streaming data processing were covered by our workloads's design:

\begin{itemize}
\item Handling parallel input streams (e.g., Kafka Stream).
\item How to aggregate the state of input stream.
\item Manage the order of stream records.
\item How to provide and update global data model in a stream processing task.
\item Evaluate the performance by measuring the latency and throughput.
\end{itemize}
We give the description of workloads, the relation with these aspects and  the implementation details of Spark's and Flink's solutions in Section 3.2.1 \& 3.2.1. Afterward, In Section 4 we discuss the performance evaluation and analysis of stream processing workloads in the two frameworks.
